Automated Topic Naming

### 2011-09-06 ese/ese.tex
66% editted

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project,
 the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly. PostgreSQL was annotated by both annotators in order
to evaluate inter-rater reliability.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?

@@

To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space considerations (but see the Appendix for our data repository). 
It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
words \emph{ease} and \emph{ownership}. In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
\emph{correctness} (182/305/748 topics) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were associated with fewer than ten topics. 
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

% EXP3
For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
$0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
PostgreSQL had an average of $0.42$ with a range of $0.31$ to $0.54$.
MaxDB had an average precision and recall of $0.16$ and $0.67$
while MySQL had $0.3$ and $0.48$. PostgreSQL had precision and recall of $0.27$ and $0.95$, respectively.

Based on this we found that \emph{reliability} and
\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
\textsf{exp3}. 
\textsf{exp1} performed poorly.
MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
but other qualities such as \emph{functionality} did improve. 
For PostgreSQL in \textsf{exp2}, \emph{reliability} and \emph{efficiency} were likewise the most accurate,
while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
Many ROC scores were $0.6$ or less, but our classifier still performed substantially better than random.

@@

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the best learner possible.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have
quite different results, as the ROC values for \emph{reliability} and
\emph{functionality} swap between projects. For PostgreSQL, the
performance is nearly always more poor than the other two systems. The
reason for this lack of performance could be that the number of
topics, $N$ could be non-optimal for PostgreSQL. Perhaps topics were
getting too mixed and it was becoming hard to distinguish one topic
from the next. PostgreSQL's dataset was the largest of the 3 projects,
the XML data-dump that described the PostgreSQL topics that we
annotated was 8X larger than MySQL and 1.7X larger than MaxDB. These
size differences arise from the number of commits, the number of files
and the verbosity of the commit descriptions.

@@

For all projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
and Multinomial Naive Bayes  are all based on Bayes' theorem and all
assume, naively, that the features supplied are independent. 
%The features we used are word counts per message. 
One beneficial aspect of this result is that it suggests we can have
very fast training and classifying  since training or classifying one
instance of Naive Bayes can be calculated in $O(N)$
for $N$ features.
%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html

@@

The less-frequently occurring a label, the harder it is to get accurate
results, due to the high noise level. Nevertheless, these results are
better than our previous word-list results of \textsf{exp2} and
\textsf{exp3}, because the ROC values are sufficiently higher in most
cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PgSQL \emph{maintainability}). The
limitation of the approach we took here is that we assume labels are
independent; however, labels could be correlated with each other. 
The next section (\ref{sec:multilabel})
addresses the issue of a lack of independence and correlation between
labels
using multi-label learners.
%In the next section we will evaluate how well these learners perform
%together.
%\DONE[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}

@@

Two perspectives to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.

@@

Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL
product. In this figure, we show the relative differences when we use
different training data sets. In \ref{fig:pgcomp-a} we see that ROC
results are very similar. In the other figure, however, author \#1 has
dramatically better performance. We speculate that this is due to the
particular annotation decisions made by author \#1; in some sense he
performed better. 
The difference in Macro-ROC was not signifcant, but the difference in
Micro-ROC was as the p-value of Student's T-test was $< 0.001$. 
%\DONE{check this writeup of comparison. T-test useful}

>>>>

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project,
 the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly. PostgreSQL (PgSQL) was annotated by both annotators in order
to evaluate inter-rater reliability.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?

@@

To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `\emph{ilities}' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space considerations (but see the Appendix for our data repository). 
It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with
words \emph{ease} and \emph{ownership}. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
\emph{correctness} (182/305/748 -- MySQL, MaxDB and Postgresql topics) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were associated with fewer than ten topics. 
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

% EXP3
For \textsf{exp3}, the average F-measure (macro-F1) for MaxDB was $0.26$ with a range $0.11$ to
$0.47$, and $0.36$ for MySQL with a range of $0.10$ to $0.65$. 
PostgreSQL had an average of $0.42$ with a range of $0.31$ to $0.54$.
MaxDB had an average precision and recall of $0.16$ and $0.67$
while MySQL had $0.3$ and $0.48$. PostgreSQL had average precision and recall of $0.27$ and $0.95$, respectively.

Based on this we found that \emph{reliability} and
\emph{usability} worked well for MaxDB in \textsf{exp2} and better in
\textsf{exp3}. 
\textsf{exp1} performed poorly.
MySQL had reasonable results within \textsf{exp2} for \emph{reliability} and \emph{efficiency}. 
MySQL's results for \emph{efficiency} did not improve in \textsf{exp3}
but other qualities such as \emph{functionality} did improve. 
For PostgreSQL in \textsf{exp2}, \emph{reliability} and \emph{efficiency} were likewise the most accurate,
while \emph{functionality} remained poor. \emph{Functionality} improved dramatically by \textsf{exp3}.
Many ROC scores were $0.6$ or less, but our classifier, in most cases, still performed substantially better than random.

@@

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the
best performing learner possible.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

Figure \ref{fig:best-learn-per-tag} shows that MaxDB and MySQL have
quite different results, as the ROC values for \emph{reliability} and
\emph{functionality} swap between projects. For PostgreSQL, the
performance is nearly always more poor than the other two systems. The
reason for this lack of performance could be that the number of
topics, $N$ could be non-optimal for PostgreSQL. Perhaps topics were
getting too mixed and it was becoming hard to distinguish one topic
from the next. PostgreSQL's dataset was the largest of the 3 projects:
the XML data-dump that described the PostgreSQL topics that we
annotated was 8 times larger than MySQL and 1.7 times larger than MaxDB. These
size differences arise from the number of commits, the number of files
and the verbosity of the commit descriptions.

@@

For all projects Bayesian techniques did the best out of a wide variety of machine learners tested. 
Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
and Multinomial Naive Bayes  are all based on Bayes' theorem and all
assume, naively, that the features supplied are independent. 
%The features we used are word counts per message. 
One beneficial aspect of this result is that it suggests we can have
very fast training and classifying  since training on or classifying one
instance with Naive Bayes can be calculated in $O(N)$
for $N$ features.
%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html

@@

The less-frequently occurring a label, the harder it is to get accurate
results, due to the high noise level. Nevertheless, these results are
better than our previous word-list results of \textsf{exp2} and
\textsf{exp3}, because the ROC values are sufficiently higher in most
cases (other than MaxDB \emph{reliability}, MySQL \emph{efficiency}, and PostgreSQL \emph{maintainability}). The
limitation of the approach we took here is that we assume labels are
independent; however, labels could be correlated with each other. 
The next section (\ref{sec:multilabel})
addresses the issue of a lack of independence and correlation between
labels
using multi-label learners.
%In the next section we will evaluate how well these learners perform
%together.
%\DONE[inline]{Comment about correlation, we should link to the Mulan section, perfect segway since Mulan relies on correlation.}

@@

Two perspectives used to evaluate multi-label learners are with micro or macro measurements (used in Figure \ref{fig:mulan}).
Macro measurements are aggregated at a class or label level (per class) while micro measurements are at the element level (per element).
A macro-ROC measurement is the average ROC over the ROC values for all labels, where a micro-ROC is the average ROC over all examples that were classified. 
For MaxDB, the macro-ROC values are undefined because of poor performance of one of the labels.%, see Figure \ref{fig:mulan}.

@@

Figure~\ref{fig:pgcomp} shows the ROC results for the PostgreSQL
product. In this figure, we show the relative differences when we use
different training data sets from different annotators. In \ref{fig:pgcomp-a} we see that ROC
results are very similar. In the other figure, however, author \#1 has
dramatically better performance. We speculate that this is due to the
particular annotation decisions made by author \#1; in some sense he
performed better. 
The difference in Macro-ROC was not significant, but the difference in
Micro-ROC was, as the p-value of Student's T-test was $< 0.001$. 
%\DONE{check this writeup of comparison. T-test useful}
